Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations

### 2021-12-07 analysis/paper/paper.Rmd
incorporated RC edits

The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* populations including Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017]; all of which have become depleted in recent decades. Declines in survival of local juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions and regional climate [@Welch2021; @Gosselin2021], and changes to the freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O. mykiss irideus* and brook trout *Salvelinus fontinalis*) among others [@IdahoOSCTeam2019]. Moreoever, the abundance of returning adults are further impacted by ocean and downriver harvests, poor ocean conditions, and changes to the spawning migration corridor. Each of these factors have contributed, to varying and unknown extents, to reduced adult escapement, the primary metric used to assess population viability. In response to the decline in Chinook Salmon abundance from the myriad human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in the rehabilitation of tributary ecosystems.

One potentially important, but perhaps under-appreciated source of mortality for Chinook Salmon is predation on emigrating juveniles by piscivorous fishes, including both native and non-native species. Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of predation on salmonids [@Petersen1994; @Ward1995]. There are generally two mechanisms that explain these high predation zones. First, migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation. Second, reservoirs and downstream tailraces created by dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species. In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of non-native Smallmouth Bass during some seasons [@Erhardt2018]. And non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River[@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020 related to their preference for slow water habitats.

In addition to non-native species, Northern Pikeminnow are a native, piscivorous fish preferring warm, slow water habitats. Consequently, they too have benefited from dams on the Columbia River and become abundant predators on salmonid outmigrants [@Knutsen1999]. Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996]. Predation by piscivorous fishes, including Northern Pikeminnow, may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021]. However, there is a dearth of current data on piscivorous predator populations in the Columbia River.

Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to dams and reservoirs. One such area is the Deadwater Slough in the Salmon River, Idaho which is an unnaturally slow and deep reach of the mainstem river approximately 1.5 kilometers in length and occupying 29.5 acres (Figure \@ref(fig:map)). The slough contains "reservoir-like" conditions which typically favor piscivorous fish predators like Northern Pikeminnow or Smallmouth Bass [@Watkins2015], and further, lacks hydrological and structural features and fish cover (e.g., large wood or substrate, woody debris, bank undercuts) that can provide essential refuge from predation for juvenile fishes. As a result, predation of juvenile Chinook Salmon and other local ESA-listed salmonids (Sockeye Salmon *O. nerka*, redband steelhead *O. mykiss gairdneri*) has been cited as a concern for populations in the Upper Salmon River

Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014]. The DSR migrants leave the natal spawning area as subyearlings between June and november and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts. Alternatively, NRR migrants remain in their natal spawning areas for approximately one year after emergence until emigration to the ocean as smolts. Diversity of migratory tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency.

The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing area for DSR emigrants. Moreover, the slough is part of the migratory pathway for not only juvenile Chinook Salmon (both DSR and NRR emigrants), but also for the endangered Snake River Sockeye Salmon population and several population of threatened Snake River steelhead. As such, the main Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants. @Axel2015 demonstrated decreased rates of emigration and survival for Sockeye Salmon from Redfish Lake during the spring emigration. Further, recent telemetry studies of Chinook salmon emigration during fall and early winter months indicated a decreased transition probability in the reach spanning Deadwater Slough (approximately 10% less than surrounding reaches).

We hypothesize that increased densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015]. In this study, we estimated the abundance of a piscivorous fish predator population in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon. Our objectives for the study were three-fold:

1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
3. Use a bioenergetics approach to estimate the total consumption of juvenile DSR and NRR Chinook salmon during defined emigration periods.

We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and how violations of some assumptions may affect overall results and inferences from the study. We also consider how consumption of juvenile Chinook Salmon by predators at Deadwater Slough might impact adult returns, hindering recovery of populations in the Upper Salmon River MPG.

@@

## Study Site

The Deadwater Slough is located approximately 6 kms downstream from the town of North Fork, Idaho (Figure \@ref(fig:map)). The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River. Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan and creating an unnaturally slow and deep section of the river resembling a small reservoir.

@@

Our goal was to estimate the abundance of potentially piscivorous predators in the Deadwater Slough using a mark-recapture approach both during the fall and spring near the peaks of the juvenile Chinook Salmon DSR and NRR emigrations, respectively. The initial intent was to sample during fall 2019 and spring 2020 to estimate predator abundances during the emigration of the 2018 brood year. However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 sampling event was delayed until fall 2020. Although this provided the benefit of an additional abundance estimates during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional sampling event was added for spring 2021. The study design was such that both the fall 2019 and fall 2020 sampling efforts included both a "mark week" and "recapture week" intended for mark-recapture models. However, due to budget constraints the spring 2021 effort only included a single "mark week"; instead, we calculated the catch-per-unit-effort (CPUE; fish per hour) for the spring 2021 effort for comparison to the fall efforts. In the end, sampling dates were as follows: fall 2019 (mark = November 12-14; recapture = November 19-21), fall 2020 (mark = October 20-23; recapture = October 27-30); spring 2021 (mark = May 18-21). The recapture events were held just one week following the mark events to reduce opportunity for immigration/emigration of predators between events i.e., to avoid violation of a closed population for mark-recapture models.

We attempted multiple capture methods during the first year including raft electrofishing, fyke netting, snorkeling, and angling to reduce selectivity, both for species and size classes. Unfortunately, some of the methods proved ineffective. For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) we lacked the appropriate motor to effectively “chase” fish given the lack of flow. Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough. And due to debris floating in the water column, there were concerns with fyke nets being pulled from their anchored locations and becoming entangled. Additionally, there were concerns with raft electrofishers and fyke nets incidentally capturing adult steelhead that are present in the mainstem Salmon River during fall and spring. In the end, angling proved to be the most effective method for capturing piscivorous predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias). From here on, our focus was on fish captured by angling, unless otherwise noted.

Our study relied heavily on volunteer anglers who were allowed to fish anywhere within Deadwater Slough; periodically or upon filling a livewell anglers would boat or hike their catch to a processing station located at the boat ramp approximately 500 m downstream from the top of the slough. For each fish, processors would record the date of capture, species, total length (TL; mm), and whether the fish was previously marked. If not previously marked, each fish was provided a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) and marks were unique for each marking day. We also recorded whether marked individuals were released back to be captured again. For release, processors attempted to distribute released individuals throughout Deadwater Slough to facilitate mixing back into the population. We also recorded the crew (person or combination of persons) that captured each fish and the start and end times for each crew were recorded to estimate total effort.

@@

To estimate the consumption of juvenile Chinook Salmon outmigrants by Northern Pikeminnow in the Deadwater Slough, we used the Fish Bioenergetics v4.0 applicatioon developed by @Deslauriers2017. The daily rate of consumption in grams for an individual Northern Pikeminnow was based on predator and prey energy densities, predator start and end weights, and water temperatures. Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017]. Prey energy densities were fixed at 21,500 J/g for juvenile Chinook salmon based on estimate from @Moss2016. To calculate the average predator start weight, we converted the average length of Northern Pikeminnow captured during our study to a weight using a weight-length formula for Northern Pikeminnow [@Parker1995] from the FSA package [@Ogle2021]. Continuous water temperature data from 2019 were used in two alternative models, on using the baseline start and end weights and the second with a 10% increase in average end weight of Northern Pikeminnow.

We ran three bioenergetic models for this study. For the first model, we chose a 78-day period from September 1 through November 17 when DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration. During this time, water temperatures exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon. This model assumed no growth for Northern Pikeminnow. The second model was run for one-year and assumed no growth for Northern Pikeminnow. The assumption of no growth for pikeminnow was included to show what may be occurring if the population of Northern Pikeminnow is stable. If growth of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimates of consumption over a full year period. To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual Chinook Salmon. This weight was an average from juvenile Chinook Salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.

@@

Over the fall time period a single Northern Pikeminnow is modeled to eat 61 g of fish. For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon. Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.

>>>>

The Upper Salmon River major population group (MPG) supports eight independent, extant spring/summer Chinook Salmon *Oncorhynchus tshawytscha* each providing important ecological, social, and economic benefits to the region. The distinct populations include Salmon River (above Redfish Lake Creek), Valley Creek, Yankee Fork Salmon River, East Fork Salmon River, Salmon River (mainstem below Redfish Lake Creek), Pahsimeroi River, Lemhi River, and North Fork Salmon River [@NOAA2017]; all of which have become depleted in recent decades. Reduction in overall population size, and more specifically, declines in survival of juvenile Chinook Salmon have been associated with hydrosystem passage [@Schaller2014], ocean conditions, regional climate [@Welch2021; @Gosselin2021], and downstream predation by native and non-native piscivorous fish [@Poe1991; @Friesen1999; @Harvey2004]. Anthropogenic changes to freshwater habitat including removal of beavers, river channel simplification, water withdrawals, and the proliferation of non-native species (e.g., non-native coastal rainbow trout *O. mykiss irideus* and brook trout *Salvelinus fontinalis*) among others [@IdahoOSCTeam2019] have also been attributed to the reductions in the population. Moreover, the abundance of returning adults is futher impacted by ocean and mainstem harvests, poor ocean conditions, and changes to the spawning migration corridor [@FCRPSBiOp2020]. Each of these factors have contributed, to varying extents, to reduced adult escapement, the primary metric used to assess population viability. In response to the decline in Upper Salmon Chinook Salmon abundance from the myriad of human activities and associated habitat degradation, action agencies have attempted to improve juvenile survival and adult spawning conditions by investing in rehabilitation of tributary ecosystems as well as investing in the study and evaluation of many factors attributed to the species decline [@Roni].

One potentially important, but perhaps under-appreciated source of mortality for Upper Salmon Chinook Salmon is predation on emigrating juveniles by piscine fishes. Research in the lower mainstem Snake and Columbia rivers has estimated that during peak emigration, predation on out-migrating salmonids has a significant negative impact (most notably in the presence on non-native piscivores) on the overall population and success of population recovery [@Fresh2003; @Winther2019]. However, these studies have focused on larger, mainstem and reservoir habitats and direct predation, but very little is known about the interaction between upstream, river dwelling piscine fishes, and their impact on salmonid recovery above reservoir systems of the Columbia and Snake Rivers [@Rubenson2020]. Adding to this, there is little consensus within the literature related to the magnitude of predation on salmonids in less altered, free flowing systems, with some suggesting lower levels of predation [@Brown1981; @Buchana1981] and others suggesting increased levels of predation can occur [@Tabor1993; @Ward1995; @Shively1996; @Zimmerman1999].

Dams and reservoirs in the Columbia River basin are the primary locations associated with high rates of predation on salmonids [@Petersen1994; @Ward1995]. There are generally two mechanisms that explain these high predation zones. First, migration rates of juvenile Chinook Salmon are reduced during reservoir passage [@Venditti2000], thereby increasing the time migrating smolts are vulnerable to predation. Second, reservoirs and downstream tailraces resulting from dams on the Columbia and Snake rivers have created favorable slow-water habitat for several species of piscivorous fish known to consume juvenile Chinook Salmon including both non-native (e.g., Smallmouth Bass *Micropterus dolomieu*, Walleye *Sander vitreus*) and native (e.g., Northern Pikeminnow *Ptychocheilus oregonensis*) species. In the Lower Granite Reservoir on the Snake River, juvenile Chinook Salmon are common prey items of non-native Smallmouth Bass during some seasons [@Erhardt2018]. And non-native Walleye are considered a potentially important source for juvenile Chinook Salmon mortality in the Snake River[@Widener2021], particularly as their abundance has significantly increased surrounding Snake River dam facilities between 2008 and 2020 related to their preference for slow water habitats. Much effort has been devoted to understanding predation related mortality throughout the hydropower system of the Lower Snake and Columbia rivers, but little is known about native and non-native predation on salmonids throughout free-flowing sections of the Columbia Basin such as the Salmon River [@Rubenson2020].

Along with non-native species, the Salmon River supports native Northern Pikeminnow, a piscine fish tolerating and thriving in relatively warm, slow water habitats. Consequently, they too have benefited from dams on the Columbia River and become abundant predators on salmonid outmigrants [@Knutsen1999]. Indeed, Northern Pikeminnow are estimated to consume 8% (16.4 million) of the estimated 200 million juvenile salmonids emigrating through the lower Snake and Columbia rivers, annually [@Beamesderfer1996]. Predation by piscivorous fishes, including Northern Pikeminnow, may have been a factor in the below-average survival of wild Chinook Salmon in 2020 [@Widener2021]. However, there is a dearth of current data on piscivorous predator populations in the Columbia River with little known about their movement, life history, or impacts upstream of the mainstem habitats [@Gadomski2001].

Habitats containing slower water velocities and other attributes that support piscivorous predators are not limited to areas near dams and reservoirs. One such example is the Deadwater Slough on the mainstem Salmon River, Idaho which is an unnaturally slow and deep reach approximately 1.5 kilometers in length and occupying 29.5 acres (Figure \@ref(fig:map)). The slough is located approximately 6 kms downstream from the town of North Fork, Idaho and contains anthropogenically created "reservoir-like" conditions which typically favor piscivorous fish predators like Northern Pikeminnow or Smallmouth Bass [@Watkins2015], and further, lacks hydrological and structural features with little to no fish cover (e.g., large wood or substrate, woody debris, bank undercuts) that can provide essential refuge from predation for juvenile fishes. The downstream end of the slough is located at the confluence of Dump Creek and the Salmon River. Around 1897, the failure of a small mining diversion reservoir in the Dump Creek drainage resulted in an erosion event that deposited substantial amounts of sediment at the confluence of the Salmon River, forming an alluvial fan and creating an unnaturally slow and deep section of the river resembling a small reservoir. As a result, predation of juvenile Chinook Salmon and other local ESA-listed salmonids (Sockeye Salmon *O. nerka*, redband steelhead *O. mykiss gairdneri*) has been cited as a concern for populations in the Upper Salmon River.

The Deadwater Slough is in a reach of the Salmon River believed to be a historically important overwinter rearing and spring emigration area for Chinook Salmon juveniles. Spring/summer Chinook Salmon in the Upper Salmon MPG are stream-type and exhibit two distinct migration tactics; downstream rearing (DSR) and natal reach rearing (NRR) [@Copeland2014]. The DSR migrants leave the natal spawning area as subyearlings between June and november and typically overwinter in downstream, mainstem habitats until the following spring when they emigrate to the ocean as smolts. Alternatively, NRR migrants remain in their natal spawning areas for approximately one year after emergence until emigration to the ocean as smolts. Diversity of migratory tactics provides a mechanism for coping with adverse conditions in freshwater rearing and migration environments and buffers against catastrophic events, thereby increasing population resiliency. Adding to its importance, the slough is part of the migratory pathway for not only juvenile Chinook Salmon (both DSR and NRR emigrants), but also for the endangered Snake River Sockeye Salmon population and several population of threatened Snake River steelhead. As such, the main Salmon River has been the target of recent radio telemetry studies to examine the downstream movement, distribution, and (apparent survival) of juvenile salmonid emigrants. @Axel2015 demonstrated decreased rates of emigration and survival for Sockeye Salmon from Redfish Lake during the spring emigration. Further, recent telemetry studies of Chinook salmon emigration during fall and early winter months indicated a decreased transition probability in the reach spanning Deadwater Slough (approximately 10% less than surrounding reaches).

We hypothesize that increased densities of piscivorous predators in the Deadwater Slough may explain the reduced survival (or apparent survival) observed for juvenile Chinook Salmon [@Ackerman2018] and Sockeye Salmon [@Axel2015]. To test this, we estimated the abundance of a piscivorous fish predator population (Northern Pikeminnow) in the Deadwater Slough and their potential impacts to juvenile salmon emigrants, focusing on DSR and NRR Chinook Salmon. Our objectives for the study were three-fold:

1. Estimate the abundance (or relative abundance) of potential predators in the Deadwater Slough during the peaks of fall (DSR) and spring (NRR) juvenile emigrations.
2. Document predation on juvenile Chinook Salmon during the emigration periods using gastric lavage.
3. Use a bioenergetics approach to estimate consumption potential (grams) of the Northern Pikeminnow population at Deadwater Slough and the percent of consumption consisting of juvenile DSR and NRR Chinook salmon during defined emigration periods.

This research is an essential first step in characterizing conditions through the Deadwater Reach and providing data necessary to discuss potential mitigation strategies (if needed), where Northern Pikeminnow removal has proven fruitful in increasing survival through the reservoir systems [@Winther2019]. We follow with a discussion of the various assumptions that went into the mark-recapture and bioenergetics models and how violations of some assumptions may affect overall results and inferences from the study. We also consider how consumption of juvenile Chinook Salmon by predators at Deadwater Slough might impact adult returns, hindering recovery of populations in the Upper Salmon River MPG.

@@

@@

Our goal was to estimate the abundance of potentially piscivorous predators in the Deadwater Slough using a mark-recapture approach both during the fall and spring near the peaks of the juvenile Chinook Salmon DSR and NRR emigrations, respectively. The initial intent was to sample during fall 2019 and spring 2020 to estimate predator abundances during the emigration of the 2018 brood year. However, due to logistical concerns during the onset of the Covid-19 pandemic, the spring 2020 sampling event was delayed until fall 2020. Although this provided the benefit of an additional abundance estimate during the DSR emigration, estimates during the NRR emigration were still lacking, and so an additional sampling event was added for spring 2021. The study design was such that the fall 2019 and fall 2020 sampling efforts included both a "mark week" and "recapture week" intended for mark-recapture models. However, due to budget constraints the spring 2021 effort only included a single "mark week"; instead, we calculated the catch-per-unit-effort (CPUE; fish per hour) for the spring 2021 sampling event for comparison to the fall efforts. In the end, sampling dates were as follows: fall 2019 (mark = November 12-14; recapture = November 19-21), fall 2020 (mark = October 20-23; recapture = October 27-30); spring 2021 (mark = May 18-21). The recapture events were held just one week following the mark events to reduce opportunity for immigration/emigration of predators between events i.e., to avoid violation of a closed population for mark-recapture models.

We attempted a multiple of capture methods during the first year including raft electrofishing, fyke netting, snorkeling, and angling to reduce selectivity, both for species and size classes. Unfortunately, some of the methods proved ineffective. For example, fish were able to elude the electrical field of the raft electrofisher due to 1) the lack of flow in Deadwater Slough and 2) our inability to "chase" fish downstream as they evaded the electricity. Snorkeling was ineffective due to the high turbidity of the mainstem Salmon River and Deadwater Slough area. And due to debris floating in the water column, there were concerns with fyke nets being pulled from their anchored locations and becoming entangled. Additionally, there were concerns with raft electrofishers and fyke nets incidentally capturing adult steelhead that are present in the mainstem Salmon River during fall and spring. Eventually, angling proved to be the most effective method for capturing piscivorous predators while minimizing potential impacts to steelhead but recognizing that angling can have both species and size selectivity (bias). From here on, we focus on fish captured by angling, unless otherwise noted.

Our study relied heavily on volunteer anglers who were allowed to fish anywhere within Deadwater Slough; periodically or upon filling a livewell anglers would boat or hike their catch to a processing station located at the boat ramp approximately 500 m downstream from the top of the slough. For each fish, processors would record the date of capture, species, total length (TL; mm), and whether the fish was previously marked. If not previously marked, each fish was provided a physical mark (e.g., hole punch of lower caudal, upper caudal, left pelvic, right pelvic) unique for each marking day. We also recorded whether marked individuals were released back to be captured again. For release, processors attempted to distribute released individuals throughout Deadwater Slough to facilitate mixing back into the population. We also recorded the crew (person or combination of persons) that captured each fish and the start and end times for each crew were recorded to estimate total effort.

@@

To estimate the consumption of juvenile Chinook Salmon outmigrants by Northern Pikeminnow in the Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RcitationFromNick]. The daily rate of consumption in grams for an individual Northern Pikeminnow was based on predator and prey energy densities, predator start and end weights, and water temperatures. Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017]. Prey energy densities were fixed at 21,500 J/g for juvenile Chinook salmon based on an estimate from @Moss2016. We used a weight-length formula [@Parker1995] to calculate the average predator start weight, converted from the average length of Northern Pikeminnow captured during our study within the FSA package [@Ogle2021]. Continuous water temperature data from 2019 was used in two alternative models, one using the baseline start and end weights and the second with a 10% increase in average end weight of Northern Pikeminnow.

We ran three bioenergetic models for this study. For the first model, we chose a 78-day period from September 1 through November 17 where DSR emigrants are known to enter the mainstem Salmon River from natal tributaries (e.g., Lemhi River) and begin their downstream migration. During this time, water temperatures exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon. This model assumed no growth for Northern Pikeminnow. The second model was run for one full year and assumed no growth for Northern Pikeminnow. The assumption of no growth for pikeminnow was included to show what may be occurring if the population of Northern Pikeminnow is stable. If growth of the population is occurring, the alternative model shows how an increase of 10% body weight changes the estimates of consumption over a full year period. To estimate the number of possible Chinook Salmon that could be consumed, we assumed a weight of 12 g per individual Chinook Salmon. This weight was an average from juvenile Chinook Salmon emigrating past six rotary screw traps operating in the Upper Salmon above Deadwater Slough including traps in the Lemhi, Pahsimeroi, and North Fork Salmon rivers and one trap operating near the Sawtooth hatchery.

@@

Over the fall time period a single average-sized Northern Pikeminnow is modeled to eat 61 g of fish. For the estimated population of ***xx,xxx***, then ***xxx,xxx*** g of fish are being eaten each fall during the migration of juvenile Chinook Salmon. Therefore, Northern Pikeminnow in the Deadwater Slough can potentially consume ***xx,xxx*** juvenile Chinook Salmon as a discrete population.
